Efficiently Handling Guards in the Low Level Design of RPython's tracing JIT

### 2012-08-06 talk/vmil2012/paper.tex
typo

After optimization the resulting trace is handed to the over platform specific
backend to be compiled to machine code. The compilation phase consists of two
passes over the lists of instructions, a backwards pass to calculate live
ranges of IR-level variables and a forward one to emit the instructions. During
the forward pass IR-level variables are assigned to registers and stack
locations by the register allocator according to the requirements of the to be
emitted instructions.  Eviction/spilling is performed based on the live range
information collected in the first pass. Each IR instruction is transformed
into one or more machine level instructions that implement the required
semantics, operations withouth side effects whose result is not used are not
emitted. Guards instructions are transformed into fast checks at the machine
code level that verify the corresponding condition.  In cases the value being
checked by the guard is not used anywhere else the guard and the operation
producing the value can merged, reducing even more the overhead of the guard.
Figure \ref{fig:trace-compiled} shows how an \texttt{int\_eq} operation
followed by a guard that checks the result of the operation are compiled to
pseudo-assembler if the operation and the guard are compiled separated or if
they are merged.

@@

Each guard in the IR has attached to it a list of the IR-variables required to
rebuild the execution state in case the trace is left through the side-exit
corresponding to the guard. When a guard is compiled, additionally to the
condition check two things are generated/compiled. First a special data
structure called \emph{low-level resume data} is created that encodes the
information provided by the register allocator about where the values
corresponding to each IR-variable required by the guard will be stored when
execution reaches the code emitted for the corresponding guard. \bivab{go into
more detail here?!} This encoding needs to be as compact as possible to
maintain an acceptable memory profile.

\todo{example for low-level resume data showing how the current encoding works?}

>>>>

After optimization the resulting trace is handed to the over platform specific
backend to be compiled to machine code. The compilation phase consists of two
passes over the lists of instructions, a backwards pass to calculate live
ranges of IR-level variables and a forward one to emit the instructions. During
the forward pass IR-level variables are assigned to registers and stack
locations by the register allocator according to the requirements of the to be
emitted instructions.  Eviction/spilling is performed based on the live range
information collected in the first pass. Each IR instruction is transformed
into one or more machine level instructions that implement the required
semantics, operations without side effects whose result is not used are not
emitted. Guards instructions are transformed into fast checks at the machine
code level that verify the corresponding condition.  In cases the value being
checked by the guard is not used anywhere else the guard and the operation
producing the value can merged, reducing even more the overhead of the guard.
Figure \ref{fig:trace-compiled} shows how an \texttt{int\_eq} operation
followed by a guard that checks the result of the operation are compiled to
pseudo-assembler if the operation and the guard are compiled separated or if
they are merged.

@@

Each guard in the IR has attached to it a list of the IR-variables required to
rebuild the execution state in case the trace is left through the side-exit
corresponding to the guard. When a guard is compiled, additionally to the
condition check two things are generated/compiled. First a special data
structure called \emph{low-level resume data} is created that encodes the
information provided by the register allocator about where the values
corresponding to each IR-variable required by the guard will be stored when
execution reaches the code emitted for the corresponding guard. This data
structure stores the data in a compressed manner using an encoding the uses
8bits to store 7bits of information. This encoding is efficient to create and
provides a compact representation of the needed information. This encoding
needs to be as compact as possible to maintain an acceptable memory profile.
